---
title: "Data Science Capstone Quiz 3"
author: "Jon Ting"
date: "24/08/2020"
output: html_document
---

# Setup Document
The knitr options and libraries needed are first loaded:
```{r Setup, warning=F, results=F}
knitr::opts_chunk$set(echo=T)
set.seed(77)
suppressPackageStartupMessages(library(tm))
```

# Prepare Datasets
The whole English dataset is rather large, consisting of more than 3 million entries in total. To reduce the time taken to implement the model, a subset of **1%** of this size is first extracted, and then split into training (80%) and validation (20%) sets. The model will be evaluated on the quiz questions so the size of the testing set is only 10:
```{r Prepare datasets, cache=T, warning=F}
# Load whole English dataset
engData <- readRDS(file="../Quiz2/eng_data.rds")
subsetProp <- 0.01
train_val_data <- sample(x=engData, size=subsetProp * length(engData))

# Train-validation split
trainProp <- 0.8
sampleNum <- length(x=train_val_data)
trainIdx <- sample(x=1:sampleNum, size=sampleNum * trainProp)
valIdx <- setdiff(x=1:sampleNum, y=trainIdx)
train <- train_val_data[trainIdx]
validation <- train_val_data[valIdx]
```

# Create Clean N-grams Data
```{r Create n-grams data, cache=T}
# Define n-gram tokenization functions
bigram <- function(x) unlist(lapply(ngrams(words(x), 2), paste, collapse=" "), use.names=F)
trigram <- function(x) unlist(lapply(ngrams(words(x), 3), paste, collapse=" "), use.names=F)
quadrigram <- function(x) unlist(lapply(ngrams(words(x), 4), paste, collapse=" "), use.names=F)
pentagram <- function(x) unlist(lapply(ngrams(words(x), 5), paste, collapse=" "), use.names=F)
hexagram <- function(x) unlist(lapply(ngrams(words(x), 6), paste, collapse=" "), use.names=F)

# Clean the corpus
delPat <- function(strng, pat) gsub(pattern=pat, replacement=" ", x=strng)
profaneWords <- read.delim(file="../Milestone/badwords.txt", header=F)[, 1]
trainCorp <- VCorpus(VectorSource(train))
trainCorp <- tm_map(x=trainCorp, FUN=content_transformer(FUN=delPat), "(f|ht)tp(s?)://(.*)[.][a-z]+")
trainCorp <- tm_map(x=trainCorp, FUN=content_transformer(FUN=delPat), "[^a-zA-Z ]") 
trainCorp <- tm_map(x=trainCorp, FUN=removePunctuation)
trainCorp <- tm_map(x=trainCorp, FUN=removeNumbers)
trainCorp <- tm_map(x=trainCorp, FUN=content_transformer(FUN=tolower))
trainCorp <- tm_map(x=trainCorp, FUN=removeWords, stopwords(kind="en"))
trainCorp <- tm_map(x=trainCorp, FUN=removeWords, profaneWords)
stemTrainCorp <- tm_map(x=trainCorp, FUN=stemDocument)
stemTrainCorp <- tm_map(x=stemTrainCorp, FUN=stripWhitespace)
stemTrainCorp <- tm_map(x=stemTrainCorp, FUN=PlainTextDocument)
unstemTrainCorp <- tm_map(x=trainCorp, FUN=stripWhitespace)
unstemTrainCorp <- tm_map(x=unstemTrainCorp, FUN=PlainTextDocument)

# Compute document term matrices for each n-gram (up to n=4)
docTermMat2 <- DocumentTermMatrix(stemTrainCorp, control=list(tokenize=bigram))
docTermMat3 <- DocumentTermMatrix(stemTrainCorp, control=list(tokenize=trigram))
docTermMat4 <- DocumentTermMatrix(stemTrainCorp, control=list(tokenize=quadrigram))
docTermMat5 <- DocumentTermMatrix(stemTrainCorp, control=list(tokenize=pentagram))
docTermMat6 <- DocumentTermMatrix(stemTrainCorp, control=list(tokenize=hexagram))

# Save the matrices
saveRDS(docTermMat2, file="./dtm2.rds")
saveRDS(docTermMat3, file="./dtm3.rds")
saveRDS(docTermMat4, file="./dtm4.rds")
saveRDS(docTermMat5, file="./dtm5.rds")
saveRDS(docTermMat6, file="./dtm6.rds")
nGramsList <- c("dtm2.rds", "dtm3.rds", "dtm4.rds", "dtm5.rds", "dtm6.rds")
```

# Implement the New Model
```{r Predictive model, cache=T}

# Apply cleaning transformations
txtTransfrom <- function(txt) {
  txt <- delPat(strng=txt, pat="(f|ht)tp(s?)://(.*)[.][a-z]+")
  txt <- delPat(strng=txt, pat="[^a-zA-Z ]") 
  txt <- removePunctuation(x=txt)
  txt <- removeNumbers(x=txt)
  txt <- tolower(x=txt)
  txt <- removeWords(x=txt, words=stopwords(kind="en"))
  txt <- removeWords(x=txt, words=profaneWords)
  txt <- stemDocument(x=txt)
  txt <- stripWhitespace(x=txt)
  return(txt)
}

# Implement backoff-model
nextWord <- function(inp_txt) { 
  
  # Extract usable parts from the input text
  inp_txt <- txtTransfrom(txt=inp_txt)
  phrase <- strsplit(x=inp_txt, split=" ")[[1]]
  if (length(phrase) == 0) { return("No text is detected after cleaning!") }
  else if ((length(unique(phrase)) == 1) & (unique(phrase) == "")) { return("Only spaces left after cleaning!") }
  else if (length(phrase) > 5) {phrase <- phrase[1:5]}
  
  # Search for entries containing the phrase of length n in stored (n+1)-grams
  searchOrder <- length(phrase):1
  for (i in seq_along(searchOrder)) {
    phrase_str <- paste(phrase[i:length(phrase)], collapse=" ")
    docTermMat <- readRDS(file=nGramsList[searchOrder[i]])
    grams <- docTermMat$dimnames[["Terms"]]
    entries <- grams[grepl(pattern=phrase_str, x=grams, ignore.case=T)]
    if (length(entries) == 0) next else break }
  if (length(entries) == 0) { return("No similar phrase found in training data, sorry!") }

  # Create a corpus from the entries containing the phrase
  regex_str <- paste(phrase_str, "([^ ]+)")
  targetWords <- ''
  for (i in 1:length(entries)) { 
    match_idx <- regexec(pattern=regex_str, text=entries[i], ignore.case=T)
    targetWords <- c(targetWords, regmatches(x=entries[i], m=match_idx)[[1]][2]) }
  targetWords <- targetWords[(!is.na(targetWords)) & (targetWords != "")]
  if (length(targetWords) == 0) { return("No similar phrase found in training data, sorry!") }
  corp <- VCorpus(VectorSource(data.frame(targetWords)))
  
  # Compare probability for each unigram and return the most likely word
  targetDTM <- as.matrix(x=DocumentTermMatrix(corp))
  freq <- sort(colSums(x=targetDTM), decreasing=T)
  df <- data.frame(Word=names(freq), Counts=freq, Probability=freq/length(freq))
  rownames(df) <- 1:length(freq)
  return(stemCompletion(x=df[1, 1], dictionary=unstemTrainCorp, type="first")[[1]])
}
```

# Estimate Computational Complexity
Using the same phrases for the previous model, this model was also timed for comparison:
```{r Time used, cache=T}
system.time(expr=nextWord("in a case of"))
system.time(expr=nextWord("a case of"))
system.time(expr=nextWord("case of"))
```

The trend of increment in complexity as the length of input text increases is clearly shown. The time taken by the model to make a prediction for the same phrases used in previous model is shorter by about 3 orders of magnitude. This indicates the success of computing and storing the n-grams outside of the model in reducing the computational complexity of the problem.

# Test the Model
The quiz questions are answered using the prediction function defined:
```{r Test model, cache=T, warning=F}
# Answer quiz questions
predictions <- ''
predictions[1] <- nextWord("I'd live and I'd")
predictions[2] <- nextWord("telling me about his")
predictions[3] <- nextWord("see arctic monkeys this")
predictions[4] <- nextWord("helps reduce your")
predictions[5] <- nextWord("to take a")
predictions[6] <- nextWord("to settle the")
predictions[7] <- nextWord("groceries in each")
predictions[8] <- nextWord("bottom to the")
predictions[9] <- nextWord("bruises from playing")
predictions[10] <- nextWord("of Adam Sandler's")
predictions

# Compute accuracy
answers <- c("die", "marital", "week end", "stress", "picture", "matter", "hand", "top", "outside", "movies")
correctNum <- 0
for (i in 1:length(predictions)) { if (predictions[i] == answers[i]) { correctNum <- correctNum + 1} }
correctNum / length(predictions)
```

# Conclusion  
1. Using a smaller list of profane words would reduce some processing time of the model.  
2. The uncertainty of the word prediction is reflected by the probability of the word returned.  
3. The n-gram backoff-model is inefficient as each n-gram has to be stored exterior of the model. This is rather inflexible and will tend to miss out long-term dependency between words in a long phrase. The model could not make use of the context in the earlier part of an overly-long (in this case, more than 6 words) sentence.   
4. Some of the most commonly missed n-grams are phrases that contain apostrophes. For instance, "I'd" in the first quiz question. This could potentially be fixed by using a larger training set.
5. Validation of the model has not been carried out.
